Compositional Certified Resource Bounds

### 2014-07-03 paper/main.tex
fixed t08

Example \emph{t08} shows the ability of the analysis to handle
negative and non-negative numbers.  Note that there are no
restrictions on the signs of $z$ and $y$.  We also see that we can
accurately track the size change of $y$ in the first loop.
Furthermore, \emph{t08} shows that we do not handle the constants $1$
or $0$ in any special way.  In all examples you could replace $0$ and
$1$ with other constants like we did in the second loop and still
derive a tight bound.  The only information, that the analyzer needs
is $y \geq c$ before assigning $y = y - c$.

@@

$3.1|[y,z]| + 0.1|[0,y]|$
\\[.7\baselineskip] 
      {\bf t08}
    \end{center}
  \end{minipage}
%
\hfill
%
%
%
  \begin{minipage}[b]{\progwidth}
    \begin{center}
   \begin{lstlisting}
while (n<0) {
  ${\color{blue}%
  n{<}0; \,  P(n,y) \vdash }$
  n=n+1;
  ${\color{blue}%
  \dashv \cdot; \,  59 {+} P(n,y) \vdash }$
  y=y+1000;
  ${\color{blue}%
  \dashv \cdot; \, 9 {+} P(n,y) \vdash }$
  while (y>=100 && *){
    ${\color{blue}%
    y{>}99; \, 9 {+} P(n,y) \vdash }$
    y=y-100;
    ${\color{blue}%
    \dashv \cdot; \, 14 {+} P(n,y) \vdash }$
    tick(5);
    ${\color{blue}%
    \dashv \cdot; \, 9 {+} P(n,y) }$
  }
  ${\color{blue}%
  \dashv \cdot; \, 9 {+} P(n,y) \vdash }$
  tick(9);
  ${\color{blue}%
  \dashv \cdot; \, P(n,y) }$
}
   \end{lstlisting}
$59|[n,0]| {+} 0.05|[0,y]|$
\\[.7\baselineskip]
      {\bf t27}
    \end{center}
  \end{minipage}

@@

\begin{figure*}
\centering
\begin{tabular}{r|c|c|c|c}
%
&
\begin{lstlisting}
while (y > x) {
  x++; tick(1); }
while (x > 2) {
  x -= 3; tick(1); }
\end{lstlisting}
&
\begin{lstlisting}
while (i > 100) {
  i--; tick(1) }
i += k+50;
while (i >= 0) {
  i--; tick(1); }
\end{lstlisting}
&
\begin{lstlisting}
while (x > 0) {
  x--;
  t=x, x=y, y=t;
  tick(1);
}
\end{lstlisting}
&
\begin{lstlisting}
assert(y >= 0);
while (x>y) {
  x -= y+1;
  for (z=y; z > 0; z--)
    tick(1);
  tick(1);
}
\end{lstlisting}
\\
%
\hline
& t08a (slightly modified t08) & t19 & t30 & t15
%
\\
\hline
LOOPUS &
{\small XXX} &
{\small $\max(0, \! i{-}10^2) \!+ \max(0, \! k {+} i {+} 51)$} &
{\small % $1 + \max(x, y)$ (???)
---
} &
---
\\
%
%
%
%
Rank &
{\small $2 {+} y {-} x (???)$ } &
{\small $54+k+i$ } &
--- &
{\small $2+2x-y$ }
\\
%
%
AAA 
  & {\small $1.33|[x,y]| {+} 0.33 |[0,x]|$ }
  & $50 {+} |\inter {-1} i| {+} |\inter 0 k|$ 
  & $|\inter 0 x| {+} |\inter 0 y|$
  & $|\inter 0 x|$
\end{tabular}
\caption{ Comparison of resource bounds derived by different tools on
  several examples with linear bounds.  AAA stands for our automatic
  amortized analysis for Clight.  The output of Rank has been manually
  simplified to fit the table.  }
\label{fig:compar}
\end{figure*}

\begin{figure}
\includegraphics[width=\linewidth]{fig/bound3d}
\caption{The automatically derived bound $1.33|[x,y]| + 0.33 |[0,x]|$
  (blue lines) and the measured runtime cost (red crosses) for example
  \emph{t08a}. For $x\ge 0$ the derived bound is tight.}
\label{fig:3d}
\end{figure}

@@

  \begin{minipage}[b]{\progwidth}
    \begin{center}
   \begin{lstlisting}
  while (z-y>0) {
    y = y+1;
  }
  while (y>9) {
    y=y-10;
  }
   \end{lstlisting}

$1.1|[y,z]| + 0.1|[0,y]|$
\\[.7\baselineskip] 
      {\bf t08}
    \end{center}
  \end{minipage}
%
%
%
  \begin{minipage}[b]{\progwidth}
    \begin{center}
   \begin{lstlisting}
  while (x-y>0) {
    if (*)
      y=y+1;
    else
      x=x-1;
  }
   \end{lstlisting}

@@

Example \emph{t08} shows the ability of the analysis to handle
negative and non-negative numbers.  Note that there are no
restrictions on the signs of $z$ and $y$.  We also see again that we
accurately track the size change of $y$ in the first loop.
Furthermore, \emph{t08} shows that we do not handle the constants $1$
or $0$ in any special way.  In all examples you could replace $0$ and
$1$ with other constants like we did in the second loop and still
derive a tight bound.  The only information, that the analyzer needs
is $y \geq c$ before assigning $y = y - c$.

@@

% $\max(0, (\max(0, (y + -9)) + \max(0, (z - y)))) + \max(0, (z - y))$ &
$\max(0, \! y{-}9)\dots$ &
$O(n)$ &

>>>>

Example \emph{t08a} shows the ability of the analysis to handle
negative and non-negative numbers.  Note that there are no
restrictions on the signs of $z$ and $y$.  We also see that we can
accurately track the size change of $y$ in the first loop.
Furthermore, \emph{t08a} shows that we do not handle the constants $1$
or $0$ in any special way.  In all examples you could replace $0$ and
$1$ with other constants like we did in the second loop and still
derive a tight bound.  The only information, that the analyzer needs
is $y \geq c$ before assigning $y = y - c$.

@@

$3.1|[y,z]| + 0.1|[0,y]|$
\\[.7\baselineskip] 
      {\bf t08a}
    \end{center}
  \end{minipage}
%
\hfill
%
%
%
  \begin{minipage}[b]{\progwidth}
    \begin{center}
   \begin{lstlisting}
while (n<0) {
  ${\color{blue}%
  n{<}0; \,  P(n,y) \vdash }$
  n=n+1;
  ${\color{blue}%
  \dashv \cdot; \,  59 {+} P(n,y) \vdash }$
  y=y+1000;
  ${\color{blue}%
  \dashv \cdot; \, 9 {+} P(n,y) \vdash }$
  while (y>=100 && *){
    ${\color{blue}%
    y{>}99; \, 9 {+} P(n,y) \vdash }$
    y=y-100;
    ${\color{blue}%
    \dashv \cdot; \, 14 {+} P(n,y) \vdash }$
    tick(5);
    ${\color{blue}%
    \dashv \cdot; \, 9 {+} P(n,y) }$
  }
  ${\color{blue}%
  \dashv \cdot; \, 9 {+} P(n,y) \vdash }$
  tick(9);
  ${\color{blue}%
  \dashv \cdot; \, P(n,y) }$
}
   \end{lstlisting}
$59|[n,0]| {+} 0.05|[0,y]|$
\\[.7\baselineskip]
      {\bf t27}
    \end{center}
  \end{minipage}

@@

\begin{figure*}
\centering
\begin{tabular}{r|c|c|c|c}
%
&
\begin{lstlisting}
while (y > x) {
  x++; tick(1); }
while (x > 2) {
  x -= 3; tick(1); }
\end{lstlisting}
&
\begin{lstlisting}
while (i > 100) {
  i--; tick(1) }
i += k+50;
while (i >= 0) {
  i--; tick(1); }
\end{lstlisting}
&
\begin{lstlisting}
while (x > 0) {
  x--;
  t=x, x=y, y=t;
  tick(1);
}
\end{lstlisting}
&
\begin{lstlisting}
assert(y >= 0);
while (x>y) {
  x -= y+1;
  for (z=y; z > 0; z--)
    tick(1);
  tick(1);
}
\end{lstlisting}
\\
%
\hline
& t08a (slightly modified t08) & t19 & t30 & t15
%
\\
\hline
LOOPUS &
{\small $\max(0, x{-}2) {+} 2\max(0, y {-} x)$} &
{\small $\max(0, \! i{-}10^2) \!+ \max(0, \! k {+} i {+} 51)$} &
{\small % $1 + \max(x, y)$ (???)
---
} &
---
\\
%
%
%
%
Rank &
{\small $2 {+} y {-} x (???)$ } &
{\small $54+k+i$ } &
--- &
{\small $2+2x-y$ }
\\
%
%
AAA 
  & {\small $1.33|[x,y]| {+} 0.33 |[0,x]|$ }
  & $50 {+} |\inter {-1} i| {+} |\inter 0 k|$ 
  & $|\inter 0 x| {+} |\inter 0 y|$
  & $|\inter 0 x|$
\end{tabular}
\caption{ Comparison of resource bounds derived by different tools on
  several examples with linear bounds.  AAA stands for our automatic
  amortized analysis for Clight.  The output of Rank has been manually
  simplified to fit the table.  }
\label{fig:compar}
\end{figure*}

\begin{figure}
\includegraphics[width=\linewidth]{fig/bound3d}
\caption{The automatically derived bound $1.33|[x,y]| + 0.33 |[0,x]|$
  (blue lines) and the measured runtime cost (red crosses) for example
  \emph{t08}. For $x\ge 0$ the derived bound is tight.}
\label{fig:3d}
\end{figure}

@@

  \begin{minipage}[b]{\progwidth}
    \begin{center}
   \begin{lstlisting}
  while (y-x>0) {
    x = x+1;
  }
  while (x>2) {
    x=x-3;
  }
   \end{lstlisting}

$1.33|[x,y]| + 0.33|[0,x]|$
\\[.7\baselineskip] 
      {\bf t08}
    \end{center}
  \end{minipage}
%
%
%
  \begin{minipage}[b]{\progwidth}
    \begin{center}
   \begin{lstlisting}
  while (x-y>0) {
    if (*)
      y=y+1;
    else
      x=x-1;
  }
   \end{lstlisting}

@@

Example \emph{t08} shows the ability of the analysis to handle
negative and non-negative numbers.  Note that there are no
restrictions on the signs of $y$ and $x$.  We also see again that we
accurately track the size change of $x$ in the first loop.
Furthermore, \emph{t08} shows that we do not handle the constants $1$
or $0$ in any special way.  In all examples you could replace $0$ and
$1$ with other constants like we did in the second loop and still
derive a tight bound.  The only information, that the analyzer needs
is $x \geq c$ before assigning $x = x - c$.

@@

% $\max(0, (\max(0, (y + -2)) + \max(0, (z - y)))) + \max(0, (z - y))$ &
$\max(0, \! y{-}2)\dots$ &
$O(n)$ &
